---
title: "Global emissions from power subsector vs GDP"
author: "Aakriti Poudel"
format:
  html:
    code-fold: show
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  error: false
---

Potential questions for the final project
1.	Which countries achieved emission reductions while maintaining GDP growth?
2.	Do countries with similar GDP levels show similar or different sub-sector emission trends?
3.	When emissions reduce in one sector, do they increase in other sub-sectors?

This dataset provides annual country-level emissions from year 2015 to 2024, disaggregated by 'power' subsector. The data was derived from Climate Trace (<https://climatetrace.org/data>). 

```{r}
# Load all necessary libraries
library(tidyverse)
library(dplyr)
library(janitor)
library(here)
```

```{r}
# Read the data for power sector
electricity <- read_csv(here::here('data', 'power', 'DATA', 'electricity-generation_country_emissions_v5_1_0.csv'),
                        na = c(" ", "0", "0.0", "NA")) %>%
  clean_names()

heat_plants <- read_csv(here::here('data', 'power', 'DATA', 'heat-plants_country_emissions_v5_1_0.csv'),
                        na = c(" ", "0", "0.0", "NA")) %>% 
  clean_names()

other_energy <- read_csv(here::here('data', 'power', 'DATA', 'other-energy-use_country_emissions_v5_1_0.csv'),
                        na = c(" ", "0", "0.0", "NA")) %>%
  clean_names()

# Read the GDP dataset
gdp <- read_csv(here::here('data', 'API_NY.GDP.MKTP.CD_DS2_en_csv_v2_280632', 'API_NY.GDP.MKTP.CD_DS2_en_csv_v2_280632.csv'),
                        na = c(" ", "NA"),
                skip = 4) %>%
  clean_names()

```
```{r}
# Combine three data sets
power <- bind_rows(electricity, heat_plants, other_energy)

# View the result
print(power)
```


```{r}
# Renaming column the name of country column

#df_rename <- df %>%
  #rename(country = iso3_country )
```

```{r}
power$start_time <- as.Date(power$start_time)
class(power$start_time)

```

```{r}
power_annual <- power %>%
  mutate(year = year(power$start_time)) 

class(power_annual$year)

```
```{r}
# Rename the column and filter year 2025
power_annual_cleaned <- power_annual %>%
  # Rename the column to match with gdp dataset
  rename(country_code = iso3_country) %>% 
  # Filter out the year 2025
  filter(year != 2025)

```


```{r}
# Convert data to long format
gdp_long<- gdp %>%
  select(-x70, -x1960:-x2014) %>% # Drop 'x70' column and from x1960 to x2014
  pivot_longer(cols = starts_with("x"),
               names_to = "year",
               values_to = "gdp_values") %>% 
  # Remove 'x' from the 'year' string
  mutate(year = stringr::str_replace(year, pattern = "x", replacement = ""), 
  # Convert the resulting string to a numeric data type
  year = as.numeric(year))
```

```{r}
# Join the data frame and reorder the columns
power_gdp <- power_annual_cleaned %>%
  left_join(gdp_long, by = c("country_code", "year")) %>% 
  # Reorder the columns by position
  select(13, 1, 12, everything())

```


```{r}
# power_mean <- power_annual %>%
#   group_by(iso3_country) %>% 
#   summarize(mean = mean(emissions_quantity)) %>% 
#   arrange(desc(mean))
```

```{r}
# # Top 10 countries for emission
# #top_ten <- list(df_mean$country[1:10])
# #class(top_ten)
# 
# top_ten <- df_mean$country[1:10]

```

```{r}
# top_countries <- df_year %>%
#   filter(country %in% top_ten)
  
```

```{r}
# plot <- ggplot(top_countries, aes(x = start_time, y = emissions_quantity, color = country)) +
#                geom_line() +
#   theme_classic() +
#   scale_x_date(
#     date_breaks = 'years',
#     date_labels = "%Y") +
#   scale_y_continuous(
#     # Divide the values by 1,000,000
#     labels = scales::label_number(scale = 1e-6, suffix = "M t")) +
#   labs(x = "Year",
#        y = "CO2 emission per metric tonnes",
#        title = "Annual top 10 carbon emitting countries from 2015 to 2024
#        by road transportation subsector")
# 
# plot
```

```{r}

# ggsave("carbon-emission.png", width = 6, height = 4)
```

```{r}
# This graph represents the annual top 10 carbon emitting countries by road transportation subsector from 2015 to 2024. The graph shows that the US is the most carbon emitting country, followed by China and Japan.
```
